---
title: "HW01"
output:
  html_document: default
  pdf_document: default
---

Students: Vivek Sharma - Quynh Le

We've worked together in pair to do this homework.


# Problem 1. 
Write a function confBand(x, y, conf=0.95) taking in a predictor vector (x1; : : : ; xn) and a response vector y = (y1; : : : ; yn) and return a plot with the points (x1; y1); : : : ; (xn; yn), the least squares line, and the confidence band at level conf. Apply your function to hp and mpg from the 04cars dataset.

```{r}
library("ggplot2")
```

```{r}

confBand <- function(x, y, conf = 0.95 ) {
  data_f = data.frame(x = x, y = y)
  model <- lm(y ~ x, data = data_f)
  model_summary = summary(model)
  p = model_summary$df[1]
  n_minus_p  = model_summary$df[2]
  F_val <- qf(conf, p, n_minus_p)
  K   = sqrt(p*F_val)
  OUT  = predict(model, newdata = data_f, se.fit = TRUE, interval = "confidence", level = conf)
  y_h = OUT$fit[,1]
  lb = y_h - K*OUT$se.fit
  ub = y_h + K*OUT$se.fit
  final_df = data.frame(lower_bound = lb, mean = y_h, upper_bound = ub, y = y, x = x)
  beta_0 = model$coefficients[1]
  beta_1 = model$coefficients[2]
  pl = ggplot(data = final_df, aes(x = x, y = y)) +
    geom_point() + 
    geom_ribbon(data = final_df, aes(ymin = lower_bound, ymax = upper_bound), alpha = 0.2, fill = 'red') + 
    geom_abline(slope = beta_1, intercept = beta_0, col = 'red')
  return (pl)
}
```


```{r}
library(ggplot2)
load("04cars.rda") # loads cars dataset"
tmp = dat[,c(13,15,16,18,19)] # extract selected variables
tmp = tmp[complete.cases(tmp),] # extracts complete cases
tmp = as.data.frame(tmp)
names(tmp) = c("hp","mpg","wt","len","wd") # abbreviate names

dat = tmp
hp = dat[,"hp"]
mpg = dat[, "mpg"]
data_f = data.frame(hp, mpg)
data_f
plt = confBand(x = data_f$hp, y = data_f$mpg, conf = 0.95)
plt
```

# Problem 2. 
Let n = 100 and draw x1; : : : ; xn iid Unif(0; 1), which stay fixed in what follows. Repeat the following experiment N = 1000 times.
- Generate yi = 1 + xi + "i, with "i i.i.d. N(0; 0:2).
- Compute the 99% confidence band and record whether it contains the true line, or not.
Summarize the result of this numerical experiment by returning the proportion of times (out of N)
that the confidence band contained the true line.

```{r}

# constants and initial params
n <-100
x <-  runif(n, min = 0, max = 1)
N <- 1000
mu <-  0; var <- 0.2; sd <- var**0.5 
cnt <- 0
cnt2 <- 0
set.seed(42)


for(i in 1:N){
  e <-  rnorm(n, mu, sd) #generate epsilon ~ Normal(0, 0.2)
  y_true = 1 + x  #True Line
  y <-  1 + x + e 
  lr <- lm(y ~ x)
  
  conf <-  confint(lr, level = 0.99)
  upper  <-  ((conf[2,2]*x) + conf[1,2])
  lower <-  ((conf[2,1]*x) + conf[1,1])
  
  # check whether the band contains true line
  temp <- data.frame(cbind(y_true, lower, upper))
  names(temp) <- c("y_true", "lower", "upper")
  outliers <- subset(temp, y_true < lower | y_true > upper)
  
  
  if (nrow(outliers) == 0){
    cnt <- cnt + 1 
  }
  else {
    cnt2 <- cnt2 + 1
  }
  
}
# returning proportion of times (out of N) that the confidence band contained the true line
print(cnt/N)


```
